hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2025-03-03 inst/manuscript/paper.Rmd
Fitting models edits

To expand the applicability of the effective value approach beyond Salmonids, **hacthR** includes a `fit_model()` function, which is species-agnostic as long as development follows a power law relationship with temperature. The function takes two input vectors--one for average incubation temperature (°C) and one for the number of days to a given phenological event--and estimates model parameters (*log~e~a* and *b*) using `stats::nls()`.

@@

Next, we generate a simulated thermal regime featuring an ascending thermograph with a mean temperature of 16 °C. Using this dataset, we apply the parameterized models for each species to predict hatch timing and total developmental duration (Figure 3). This example highlights the flexibility of **hatchR** im accommodating diverse fish species and environmental conditions, making it a valuable tool for researchers and managers working outside of Salmonid system.

@@

Note the *R^2^* fit from the models below. You can see they generally preform well and are in line with values from model 2 of @beacham1990.

>>>>

To expand the applicability of the effective value approach beyond Salmonids, **hatchR** includes a `fit_model()` function, which is species-agnostic as long as development follows a power law relationship with temperature. The function takes two input vectors--one for average incubation temperature (°C) and one for the number of days to a given phenological event--and estimates model parameters (*log~e~a* and *b*) using `stats::nls()`.

@@

To demonstrate parameterizations, we generate a simulated thermal regime featuring an ascending thermograph with a mean temperature of 16 °C (again, available in paper.Rmd). Using this dataset, we apply the parameterized models for each species to predict hatch timing and total developmental duration (Figure 3). This example highlights the flexibility of **hatchR** im accommodating diverse fish species and environmental conditions, making it a valuable tool for researchers and managers working outside of Salmonid system.

@@

Note the *R^2^* fit from the models below. You can see they generally preform well and are close to values from model 2 of @beacham1990, which fall between 0.95-0.99 range.
